Collaboration in Ad Hoc Teamwork: \\ Ambiguous Tasks, Roles, and Communication

### 2016-02-07 paper/adhoccom.tex
integrated Peter comments

\maketitle
\begin{abstract}
% Autonomous agents are becoming increasingly present in our environments, and will increasingly have to collaborate with previously unseen agents.
Creating autonomous agents capable of cooperating with previously unfamiliar teammates, known as ``ad hoc teamwork'', has been identified as an important challenge for multiagent systems. Previous research has assumed that either the task, the role of each agent, or the communication protocol among agents is known before the interaction begins. We consider these three variables simultaneously and show how an ad hoc agent can fit into a new team while handling ambiguous tasks, roles, and communication protocols. We assume a known distribution of possible tasks, roles, and communication protocols. We present experimental results in the pursuit domain showing that our ad hoc agent can join such a team without barely impacting the overall performance compared to a pre-coordinated agent.
\end{abstract}

@@

Work partially supported by INRIA, Conseil R\'{e}gional d'Aquitaine, the ERC grant EXPLORERS 24007, and a INRIA Explorer fellowship. A portion of this work has taken place in the Learning Agents Research Group (LARG) at the Artificial Intelligence Laboratory, The University of Texas at Austin.  LARG research is supported in part by grants from the National Science Foundation (CNS-1330072, CNS-1305287), ONR (21C184-01), AFRL (FA8750-14-1-0070), AFOSR (FA9550-14-1-0087), and Yujin Robot.

>>>>

\maketitle
\begin{abstract}
% Autonomous agents are becoming increasingly present in our environments, and will increasingly have to collaborate with previously unseen agents.
Creating autonomous agents capable of cooperating with previously unfamiliar teammates, known as ``ad hoc teamwork'', has been identified as an important challenge for multiagent systems. Previous research has assumed that either the task, the role of each agent, or the communication protocol among agents is known before the interaction begins. We consider these three variables simultaneously and show how an ad hoc agent can fit into a new team while handling ambiguous tasks, roles, and communication protocols. We assume a known distribution of possible tasks, roles, and communication protocols. We present experimental results in the pursuit domain showing that our ad hoc agent can join such a team while barely impacting the overall performance compared to a pre-coordinated agent.
\end{abstract}

@@

Work partially supported by INRIA, Conseil R\'{e}gional d'Aquitaine, the ERC grant EXPLORERS 24007, and a INRIA Explorer fellowship. A portion of this work has taken place in the Learning Agents Research Group (LARG) at the Artificial Intelligence Laboratory, The University of Texas at Austin.  LARG research is supported in part by grants from the National Science Foundation (CNS-1330072, CNS-1305287), ONR (21C184-01), AFRL (FA8750-14-1-0070), and AFOSR (FA9550-14-1-0087).

### 2016-02-07 paper/domain.tex
integrated Peter comments

In the remainder of this section, we describe how agents plan their actions, as well as the strategy of the predators to lock the prey at the capture state. We first assume predators have full observability of the domain and later remove this ability and describe the communication systems.

@@

For a full team, the probability map of the prey state will never be uniform, i.e.  merging the information from messages of different agents will always be coherent. As we will see in the next section, this might not be the case when the ad hoc agent tries to understand what is going on by interpreting messages according to different hypotheses on the communication protocol. Observing discrepency between messages will thus be valuable to inferring the team communication system.% As described in previous subsection, given a probability map of prey state, each predator computes a reward map given its specific role in the team (N/S/E/W), and selects the optimal action to maximize this reward.

>>>>

In the remainder of this section, we describe how agents plan their actions, as well as the strategy of the predators to surround the prey at the capture state. We first assume predators have full observability of the domain and later remove this ability and describe the communication systems.

@@

For a full team, the probability map of the prey state will never be uniform, i.e.  merging the information from messages of different agents will always be coherent. As we will see in the next section, this might not be the case when the ad hoc agent tries to understand what is going on by interpreting messages according to different hypotheses on the communication protocol. Observing a discrepency between messages will thus be valuable to inferring the team communication system.% As described in previous subsection, given a probability map of prey state, each predator computes a reward map given its specific role in the team (N/S/E/W), and selects the optimal action to maximize this reward.

### 2016-02-07 paper/introduction.tex
integrated Peter comments

A similar problem, where both the task and the communication are unknown, has been investigated in human-machine interaction \cite{grizou2014interactive}. They consider only two agents, a teacher and a learner. Their learning agent can only act on its own towards the success of the task and can observe non-symbolic communication signals from the teaching agent, whose meaning is part of finite set but initially unknown.

>>>>

A similar problem, where both the task and the communication are unknown, has been investigated in human-machine interaction \cite{grizou2014interactive}. They consider only two agents, a teacher and a learner. Their learning agent can only act on its own towards the success of the task and can observe non-symbolic communication signals from the teaching agent, whose meaning is part of a finite set but initially unknown. This work differs mainly by the multiagent scenario. On the one hand, it makes the problem less tractable, but on the other hand, it simplifies the problem because the learning agent has access to more information (i.e. it can observe other agent's actions).

### 2016-02-07 paper/problem.tex
integrated Peter comments

A domain is defined by $d = \{\env, \task, \config, \com\}$ that is a subset of all possible domains $D$. We denote $S$ as the set of all agent states, $ S~=~\{ s_{b_1}, \ldots, s_{b_{n_B}} \} $, respectively $S'$ the set of all agent next states. We denote $M$ as the set of all agents' messages, $M~=~\{m_{b_1}, \ldots, m_{b_{n_B}}\}$. We want to evaluate how an ad hoc agent $a$ can adapt in such a domain. To evaluate its performance, we remove one agent randomly from a fully formed team, creating the set $B^-$, and replace it by the ad hoc agent. The resulting team is denoted as $B^{-}_{a}$.
%
The team performance is evaluated on the task $\task$ using the reward function $R$. We denote $score(B, d)$ as the score resulting from the team $B$ executing the problem $d$, i.e. the accumulated reward. In this work, we want to create an ad hoc agent that minimizes the score loss between the original team $score(B, d)$ and the team with the ad hoc agent $score(B^{-}_{a}, d)$. The problem is that the ad hoc agent needs to fit into a team yet unknown to it. It must therefore identify all the components of the domain ($\env, \task, \config, \com$). The main challenge is that the ad hoc agent does not have direct access to its performance. Indeed, it cannot compute $score(B^{-}_{a}, d)$ because $d$ is unknown to it.

@@

To tackle this problem, we assume the agent has access to a bigger set $D = \{d_1, \ldots, d_{n_H}\}$, containing $n_H$ possible domains, from which is pulled the particular domain $d$ considered. We further consider that, for any given $d_h$, the ad hoc agent can predict, in a probabilistic way, the expected behavior and communication of the agents. Hence, our approach relies on computing the posterior probability of each hypothetical domain given the information available to the ad hoc agent, here the observation from states and messages of the other agents. The correct hypothesis will be the one that maximizes this probability:
%
\begin{eqnarray}
\argmax_h~~p(d_h|S',S,M)  \label{eq:problem}
\end{eqnarray}
%
where $S$ and $S'$ are the observed states and next states, and $M$ is the messages sent by each agent. At each step a new tuple $\{S', S, M\}$ is observed and the probabilities are updated. Following Bayes' rule, we have to compute two different components: first the probability of the observed next states given the initial states, the messages, and a domain hypothesis $p(S'|S,M,d_h)$, and then the probability of the messages themselves given the oberved states and a domain hypothesis $p(M|S,d_h)$. In the following subsections, we detail these components as well as how the ad hoc agent plans its actions.% first how to use observations of the state of other agents, then how to exploit their messages, and finally

@@

We now consider the action selection method for the ad hoc agent. Previous work considered ad hoc agents that solve the optimal teammate problem. Such agents know the model of their teammates and select their next action in order to improve maximally the team performance -- often resulting in better performance than the initial team \cite{barrett2011empirical}. The aim of this study is different; we want to advocate the fact that the ad hoc agent can work under less assumptions than before and be able to estimate more information about the new team. Hence, we isolate our algorithm from the planning aspects and test whether it can select between candidate domains. Therefore, in this work, the ad hoc agent simply tries to replace a missing agent in the team. To this end, the ad hoc agent will weight the policies for each domain hypotheses $d_h$ by the probability currently assigned to this configuration $p(d_h)$.
%
\begin{eqnarray}
p(u_a|M, S) = \sum_h p(u_a|M, S, d_h) p(d_h)
\label{eq:adhocpolicy}
\end{eqnarray}
%
With this planning strategy, once the correct hypothesis is identified, the ad hoc agent will mimic the default behavior of the agent it replaces. But the agent is likely to make sensible decision earlier as irrelevant hypotheses are discarded.

>>>>

A domain is defined by $d$$=$$\{\env, \task, \config, \com\}$ that is a subset of all possible domains $D$. We denote $S$ as the set of all agent states, $ S$$=$$\{ s_{b_1}, \ldots, s_{b_{n_B}} \} $ and $S'$ the set of all agent next states. We denote $M$ as the set of all agents' messages, $M$$=$$\{m_{b_1}, \ldots, m_{b_{n_B}}\}$. We want to evaluate how an ad hoc agent $a$ can adapt in such a domain. To evaluate its performance, we remove one agent randomly from a fully formed team, creating the set $B^-$, and replace it by the ad hoc agent. The resulting team is denoted as $B^{-}_{a}$.
%
The team performance is evaluated on the task $\task$ using the reward function $R$. We denote $score(B, d)$ as the score resulting from the team $B$ executing the problem $d$, i.e. the accumulated reward. In this work, we want to create an ad hoc agent that minimizes the score loss between the original team $score(B, d)$ and the team with the ad hoc agent $score(B^{-}_{a}, d)$. The problem is that the ad hoc agent needs to fit into a team yet unknown to it. It must therefore identify all the components of the domain ($\env, \task, \config, \com$). The main challenge is that the ad hoc agent does not have direct access to its performance. Indeed, it cannot compute $score(B^{-}_{a}, d)$ because $d$ is unknown to it.

@@

To tackle this problem, we assume the agent has access to a bigger set $D = \{d_1, \ldots, d_{n_H}\}$, containing $n_H$ possible domains, from which is pulled the particular domain $d$ considered. We further consider that, for any given $d_h$, the ad hoc agent can predict, in a probabilistic way, the expected behavior and communication of the agents. Hence, our approach relies on computing the posterior probability of each hypothetical domain given the information available to the ad hoc agent, here the observation from states and messages of the other agents. The correct hypothesis will be the one that maximizes this probability:
%
\begin{eqnarray}
\argmax_h~~p(d_h|S',S,M)  \label{eq:problem}
\end{eqnarray}
%
where $S$ and $S'$ are the observed states and next states, and $M$ is the messages sent by each agent. At each step a new tuple $(S', S, M)$ is observed and the probabilities are updated. Following Bayes' rule, we have to compute two different components: first the probability of the observed next states given the initial states, the messages, and a domain hypothesis $p(S'|S,M,d_h)$, and then the probability of the messages themselves given the oberved states and a domain hypothesis $p(M|S,d_h)$. In the following subsections, we detail these components as well as how the ad hoc agent plans its actions.% first how to use observations of the state of other agents, then how to exploit their messages, and finally

@@

We now consider the action selection method for the ad hoc agent. Previous work considered ad hoc agents that solve the optimal teammate problem. Such agents know the model of their teammates and select their next action in order to improve maximally the team performance -- often resulting in better performance than the initial team \cite{barrett2011empirical}. The aim of this study is different; we want to demonstrate the fact that the ad hoc agent can work under fewer assumptions than before and be able to estimate more information about the new team. Hence, we isolate our algorithm from the planning aspects and test whether it can select between candidate domains. Therefore, in this work, the ad hoc agent simply tries to replace a missing agent in the team. To this end, the ad hoc agent will weight the policies for each domain hypotheses $d_h$ by the probability currently assigned to this configuration $p(d_h)$.
%
\begin{eqnarray}
p(u_a|M, S) = \sum_h p(u_a|M, S, d_h) p(d_h)
\label{eq:adhocpolicy}
\end{eqnarray}
%
With this planning strategy, once the correct hypothesis is identified, the ad hoc agent will mimic the default behavior of the agent it replaces. But the agent is likely to make sensible decisions earlier as irrelevant hypotheses are discarded.

### 2016-02-07 paper/results.tex
integrated Peter comments

We now present several experiments to evaluate how an ad hoc agent can join a team for which it does not know the specific task, its role, and the communication signals being used. We will compare several teams: a pre-formed team (\emph{T}), a team including the ad hoc agent (\emph{A}), and a few baseline described next. We consider several different conditions that affect the team efficiency and the difficulty for the ad hoc agent to join the team: full observability (\emph{FO}) and partial observability without (\emph{PO}) and with (\emph{POC}) communication. We present results with 20 percent of noise in the action and communication as described in Section~\ref{sec:domain}.

@@

\begin{figure}[htbp!]
  \centering
  \includegraphics[trim=2.6cm 0.4cm 2.7cm 1.8cm, clip=true, width=0.8\columnwidth]{plots/with_noise/teamComparaison.png}
  \caption{Comparison of teams with full observability (\emph{T-FO}), partial observability without (\emph{T-PO}) and with (\emph{T-POC}) communication. The use of communication in the partial observability case allows recovering similar performances than with full observability.}
  \label{fig:cmpteam}
\end{figure}

@@

We now remove one of the agents from the standard team and replace it with our ad hoc agent (Figure~\ref{fig:fullobs}). In the case of full observability the inclusion of the ad hoc agent has no impact, on average, on the team performance (\emph{T-FO} vs \emph{A-FO} -- the null hypothesis cannot be rejected with {\small$p=0.276$}). It means that the ad hoc agent can correctly identify the correct team configuration without impacting the behavior of the full team. As a point of comparison, we added the performance of a team with one of the agents acting randomly (\emph{R-FO}). Such team almost never captures the prey.

>>>>

We now present several experiments to evaluate how an ad hoc agent can join a team for which it does not know the specific task, its role, and the communication signals being used. We will compare several teams: a pre-formed team (\emph{T}), a team including the ad hoc agent (\emph{A}), and a few baselines described next. We consider several different conditions that affect the team efficiency and the difficulty for the ad hoc agent to join the team: full observability (\emph{FO}) and partial observability without (\emph{PO}) and with (\emph{POC}) communication. We present results with 20 percent noise in the action and communication as described in Section~\ref{sec:domain}.

@@

\begin{figure}[htbp!]
  \centering
  \includegraphics[trim=2.6cm 0.4cm 2.7cm 1.8cm, clip=true, width=0.8\columnwidth]{plots/with_noise/teamComparaison.png}
  \caption{Comparison of teams with full observability (\emph{T-FO}), partial observability without (\emph{T-PO}) and with (\emph{T-POC}) communication. The use of communication in the partial observability case allows recovering similar performances to full observability.}
  \label{fig:cmpteam}
\end{figure}

@@

We now remove one of the agents from the standard team and replace it with our ad hoc agent (Figure~\ref{fig:fullobs}). In the case of full observability the inclusion of the ad hoc agent has no impact, on average, on the team performance (\emph{T-FO} vs \emph{A-FO} -- the null hypothesis cannot be rejected with {\small$p=0.276$}). It means that the ad hoc agent can correctly identify the correct team configuration without impacting the behavior of the full team. As a point of comparison, we added the performance of a team with one of the agents acting randomly (\emph{R-FO}). Such a team almost never captures the prey.
